Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators

### 2025-10-24 thesis/paper_draft.qmd
final check + changes

A collection of time series connected via a set of linear constraints is known as hierarchical time series. Forecasting these series without respecting the hierarchical nature of the data can lead to incoherent forecasts across aggregation levels and, in practice, reduced accuracy. Forecast reconciliation corrects this by adjusting base forecasts to satisfy such constraints. Among modern reconciliation methods, Minimum Trace (MinT) is widely used, however, it requires a good estimate of the forecast error covariance matrix. The current practice uses linear shrinkage towards a diagonal target. Furthermore, the covariance estimate is based on in-sample 1-step-ahead forecast errors, then proportionally scale it to approximate $h$-step-ahead covariance matrix. This leaves a question of whether this method is appropriate for all real-world applications. We study the shortcomings of current practice and assess the impacts of alternative covariance estimators on MinT, including the NOVELIST estimator (shrinkage towards a soft-thresholded target), PC-adjusted shrinkage (which utilises latent factor structures), and horizon-specific estimators that relax proportional scaling. We evaluate the accuracy of MinT using these covariance estimates for both point and probabilistic reconciled forecasts, and demonstrate their effectiveness and improvements over the shrinkage estimator in a complex, large-hierarchical dataset.

@@

I would like to express my sincere gratitude to my supervisors, Prof. George Athanasopoulos and Dr. Shanika Wickramasuriya, for their invaluable guidance, support, and encouragement throughout my Honours research journey. Their directions and insights have been instrumental not only in shaping this thesis but also in my personal and academic growth, letting me understand what it means to be a researcher. Despite being busy with their own work, they always made time to assist me whenever I knocked on their doors. I also thank for the laughs and light-hearted moments during our weekly meetings, which made the process enjoyable.

I wish to thank Assoc. Prof. Natalia Bailey for being a wonderful Honours coordinator, always ready to help and provide constructive comments after each milestone, and for organising the Honours seminar series throughout the year. She also provided great advice when I was considering pursuing Honours initially, and I would say I made the right choice. My gratitude should also go to Prof. Jiti Gao for his advice on my decision to pursue Honours.

I am grateful to my friends and peers, and to the professors and lecturers for their insightful classes throughout my Honours studies. I would also like to thank the Monash University Department of Econometrics and Business Statistics for providing a supportive and stimulating academic environment.

@@

In time series forecasting, aggregation occurs in a variety of settings. For example, Starbucks Corporation operates in many countries, and each country has multiple cities where they have outlets. The sales data is structured *hierarchically*: the top level is the company's total sales, which disaggregates into sales by country, then sales by city within each country, and finally down to sales by individual outlet within each city. As a result, there are over 50,000 sales series across all *aggregation levels*, and decision makers need forecasts at each level to manage inventory and plan marketing strategies effectively. The hierarchy can be even more complex if we consider the sales of different categories of products (e.g., beverage, food, etc.) and the sales of product within each category (e.g., latte, cappuccino, etc.) at each aggregation level. In this case, the structure is called a *grouped structure*, where the aggregation paths are not unique. Such structures also arises in many other decision-making contexts, from supply chains [@Seaman2022-bb; @Angam2025-od] and energy planning [@Di-Modica2021-ad], to macroeconomics [@El-Gemayel2022-hx; @Li2019-vv] and tourism analysis [@Athanasopoulos2009-lp]. Stakeholders in these settings need forecasts at several aggregation levels to allocate resources and manage risk.

In practice, when forecasts are produced for all series (often called *base forecasts*), they typically violate the aggregation constraints observed in the data (e.g., the sum of all countries' sales forecasts does not equal the total sales forecast). Such forecasts are called *incoherent*. Incoherence undermines downstream decisions that require internal consistency and can degrade forecasting performance. To tackle this problem, forecast reconciliation was introduced. Forecast reconciliation, a post-processing step, utilises the information from the hierarchical structure and data to adjust the initially produced base forecasts, so that the resulting *reconciled forecasts* are *coherent* (i.e., respecting the aggregation constraints). It was first introduced by @Hyndman2011-jv, and later developed by @Van_Erven2015-nx, @Hyndman2016-ic, @Ben-Taieb2019-yx, @Wickramasuriya2019-xq, @Wickramasuriya2020-uk, and others that focus point forecast reconciliation (more accurately, reconciling the means of predictive distributions). Recognising the important of probabilistic forecasting, probabilistic forecast reconciliation was studied by @Shang2017-ux, @Jeon2019-wm, and @Ben-Taieb2021-bn, and later formally defined by @Panagiotelis2023-fs. @Athanasopoulos2024-as provide a comprehensive review of the literature on forecast reconciliation.

Among the modern methods, the Min Trace (MinT) approach developed by @Wickramasuriya2019-xq is is widely used due to its strong theoretical properties for minimising total reconciled forecast error variance, computational efficiency, and robust empirical performance. MinT is later extended to probabilistic reconciliation by @Wickramasuriya2024-mb, showing that it also minimises the negative log score of the reconciled distribution under Gaussian assumptions. @Wickramasuriya2019-xq also argued that modelling spatial autocorrelations directly from the start would be challenging as in this case of a large collection of time series. Post-processing reconciliation has the advantage to implicitly model this spatial autocorrelation structure, especially true for MinT. MinT is implemented in popular R and Python software ecosystems [@O-Hara-Wild2024-we; @Nixtla].

@@

Meanwhile, there has been substantial progress in high-dimensional covariance estimation that can be leveraged for MinT. Building on shrinkage, @Huang2019-ua proposed NOVELIST, which shrinks toward a soft-thresholded target rather than a diagonal one, improving flexibility to sparse targets. Extending beyond linear shrinkage, @Ledoit2012-ga (and further developed by @Ledoit2020-xt) introduced nonlinear shrinkage that replaces sample eigenvalues with data-driven nonlinear transformation of themselves. In a complementary direction, @Fan2013-jz introduce factor-based estimators that explicitly preserve latent low-rank structure while thresholding the idiosyncratic component. Related thresholding approaches, including hard thresholding [@Bickel2008-nh], generalised thresholding [@Rothman2009-yf], and adaptive thresholding [@Cai2011-jf], have also been developed. This growing toolkit provides multiple pathways to maximise the performance of MinT.

@@

When setting $\boldsymbol{M} = \boldsymbol{I}_n$, the identity matrix, the $\boldsymbol{P}$ matrix reduces to the OLS reconciliation, which corresponds to an orthogonal projection onto $\mathfrak{s}$. For a general positive definite matrix $\boldsymbol{M}$, $\boldsymbol{P}$ is idempotent but not necessarily symmetric, resulting in an oblique projection. Thus, the choice of $\boldsymbol{M}$ determines the projection direction and, in turn, how disagreements among levels are resolved.

@@

This projection view extends seamlessly from point to probabilistic forecast reconciliation. Given an $h$-step-ahead base forecast density with mean $\hat{\boldsymbol{y}}_{t+h|t}$ and covariance $\boldsymbol{W}_h$, via the linear projection with $\boldsymbol{P}$, the resulting reconciled forecast density has mean $\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{P} \hat{\boldsymbol{y}}_{t+h|t}$ and covariance $\boldsymbol{P} \boldsymbol{W}_h \boldsymbol{P}'$. @fig-geometry schematically illustrates this projection. Reconciliation takes the (possibly elliptical) base forecast distribution and "pushes" the entire mass linearly onto the coherent subspace, resulting in the reconciled forecast distribution.

@@

@Wickramasuriya2019-xq showed that by setting $\boldsymbol{M} = \boldsymbol{W}_h = \mathbb{E}( \hat{\boldsymbol{e}}_{t+h|t} \; \hat{\boldsymbol{e}}_{t+h|t}' )$, the covariance matrix of the $h$-step-ahead base forecast errors $\hat{\boldsymbol{e}}_{t+h|t} = \boldsymbol{y}_{t+h} - \hat{\boldsymbol{y}}_{t+h|t}$, we essentially minimise the total variance of the reconciled forecast errors across all series. Equivalently, MinT is the unique linear-unbiased reconciler that minimises the trace of $\text{Var}[\boldsymbol{y}_{t+h} - \tilde{\boldsymbol{y}}_{t+h|t}] = \boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}'$. This method is thus called Minimum Trace (MinT) reconciliation. The matrix $\boldsymbol{G}_h$ is thus given by:

@@

where, $\hat{\boldsymbol{R}}_{1,\delta}$ is the thresholded correlation matrix, where each element is regularised by:

@@

On the other hand, the optimal threshold $\hat \delta$ does not have a closed-form solution, and is obtained by rolling-window cross-validation procedure. The idea is to find the threshold $\delta^*$, with the corresponding $\lambda^*$ and $\hat{\boldsymbol{R}}^{N}_{1}(\delta^*, \lambda^*)$, that minimises the average out-of-sample 1-step-ahead mean squared reconciled forecast errors over all windows. The formal algorithm is given in Algorithm \ref{alg:novelist_cv}.

@@

\For{$i = v:T-1$}
    \State $j = i - v +1$
    \State $\hat{\boldsymbol{W}}_1 = \frac{1}{v} \sum_{t=j}^{i} \hat{\boldsymbol{e}}_{t} \hat{\boldsymbol{e}}_{t}'$
    \State $\hat{\boldsymbol{D}} = \text{diag}(\hat{\boldsymbol{W}}_1)$
    \State $\hat{\boldsymbol{R}}_1 = \hat{\boldsymbol{D}}^{-1/2} \hat{\boldsymbol{W}}_1 \hat{\boldsymbol{D}}^{-1/2}$
    \For{$\delta \in \Delta$}
        \State Compute thresholded correlation $\hat{\boldsymbol{R}}_{1,\delta}$ using Equation 2
        \State Compute shrinkage parameter $\hat{\lambda}_{\delta}$ using Equation 3
        \State Compute $\hat{\boldsymbol{R}}^{N}_{1,\delta}$ using Equation 1
        \State $\hat{\boldsymbol{W}}^{N}_{1,\delta} = \hat{\boldsymbol{D}}^{1/2} \hat{\boldsymbol{R}}^{N}_{1,\delta} \hat{\boldsymbol{D}}^{1/2}$
        \State $\boldsymbol{G} = (\boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{1,\delta} \boldsymbol{S})^{-1} \boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{1,\delta}$
        \State Reconciled forecasts $\tilde{\boldsymbol{y}}_{i+1} = \boldsymbol{S} \boldsymbol{G} \hat{\boldsymbol{y}}_{i+1}$
        \State $\tilde{\boldsymbol{e}}_{i+1, \delta} = \boldsymbol{y}_{i+1} - \tilde{\boldsymbol{y}}_{i+1}$
    \EndFor
\EndFor

\State $\text{MSE}_{\delta} = \frac{1}{T-v} \sum_{i=v}^{T-1} (\tilde{\boldsymbol{e}}_{i+1, \delta})^2$ for each $\delta \in \Delta$
\State $\hat{\delta} = \arg\min_{\delta \in \Delta} \text{MSE}_{\delta}$

@@

The general design of data generating process for bottom-level series is a stationary VAR(1) process, with the following structure:

@@

These steps result in a positive covariance matrix $\boldsymbol{\Sigma}^+$ where all the elements are positive. To allow a mixture of positive and negative covariance, we randomly flip the signs of them but need to maintain the positive definiteness. This can be done by pre- and post-multiplying $\boldsymbol{\Sigma}$ by a random diagonal matrix $\boldsymbol{V}$ with diagonal entries sampled from $\{-1, 1\}$ with equal probability. The final covariance matrix is given by $\boldsymbol{\Sigma} = \boldsymbol{V} \boldsymbol{\Sigma}^+ \boldsymbol{V}$.

For all hierarchies in our experiments, we simulate two panel lengths, $T=54$ ("short") and $T=304$ ("long"), reserving the final four observations as an out-of-sample test set. We perform 500 Monte Carlo replications for each configuration. In each replication, we fit univariate ARIMA models to the training observations using the automatic AICc minimization algorithm of @Hyndman2008-rf, implemented in the *fabletools* package [@O-Hara-Wild2024-we], generating 1–4-step base forecasts. We then reconcile the base forecasts using different reconciliation methods, and evaluate their point and probabilistic forecast accuracy on the test set.

@@

In our main experiments, we examine how MinT combined with the different estimators perform as the hierarchy expands. We generate synthetic data from the VAR(1) framework described earlier, varying the number of bottom‐level series, $n_b$, across two structures: a "small" structure with six groups of six bottom-level series ($n_b = 6$x$6 = 36$), and a "large" configuration with two groups of fifty bottom-level series ($n_b = 2$x$50 = 100$).

@@

The first key observation is that methods with shrinkage slightly outperform those with NOVELIST across all scenarios, although the differences are small. Second, the PC-adjusted variants (using one and two principal components) do not yield improvements over the vanilla versions. This is expected since the synthetic data generating process does not simulate from strong latent factors. Lastly, the scaled variance and $h$-step-ahead residual approaches do not enhance performance as the forecast horizon increases, suggesting that the proportionality assumption may not be severely violated in this VAR(1) setup.

@@

## Other Data Generating Processes {#sec-sim-other}

@@

These findings motivates our turn to empirical data in the next section, where latent structural features, regime shifts, and noisy, intermittent series may possibly reveal performance differences.

@@

Forecasting domestic tourism flows is of great economic and policy significance, as policymakers and stakeholders rely on accurate forecasts to make informed decisions regarding resource allocation, infrastructure development, and marketing strategies. The domestic tourism flows in Australia exhibit a natural hierarchical and grouped structure, driven both by geography and by purpose of travel. At the top of this hierarchy lies the national total, which splits into the seven states and territories. Each state is further sub-divided into tourism zones, which in turn break down into 77 regions. A complete illustration of this geographic hierarchy appears in Appendix @sec-tourism-supplement. Intersecting this geographic hierarchy is a second dimension: policymakers are also interested in travel motives. This partitions tourism flows into four categories: holiday, business, visiting friends and relatives, and other. Altogether, this yields a grouped structure of 560 series, from the most disaggregated regional-purpose cells up to the full national aggregate. @tbl-tourism-structure depicts this structure.

@@

We quantify tourism demand via "visitor nights", the total number of nights spent by Australians away from home. The data is collected via the National Visitor Survey, managed by Tourism Research Australia, using computer assisted telephone interviews from nearly 120,000 Australian residents aged 15 years and over [@tourism-au].

The data are monthly time series spanning from January 1998 to December 2016, resulting in 228 observations per series. This gives a challenging high-dimensional forecasting problem with number of series (560) doubled the number of observations, which is ideal for evaluating reconciliation approaches that rely on high-dimensional covariance estimation. The extreme dimensionality over sample size mirrors many contemporary business problems, for instance, Starbucks Corporation sales. Tourism demand is also economically vital yet highly volatile, with geographical and purpose‑specific patterns create a realistic stress‑test for reconciliation algorithms.

@@

To assess forecasting performance between models, we adopt a rolling-window cross-validation scheme. Beginning with the first 120 monthly observations (January 1998-December 2005) as the initial training set, we obtain the best-fitted ARIMA model for each of the 560 series via the automatic algorithm by minimising AICc from @Hyndman2008-rf, implemented in the *fabletools* package [@O-Hara-Wild2024-we]. The 1- to 12-step-ahead base forecasts are then generated by these ARIMA models, and then reconciled using multiple approaches. To estimate the NOVELIST and its variants, we implement an extra cross-validation prodecure within this training window, as described in Algorithm \ref{alg:novelist_cv}. We then roll the training window forward by one month and refit all models, re-estimate reconciliation variants, and produce another batch of 1- to 12-step-ahead forecasts, repeating until the training set reaches December 2015. In total, this results in 97 out-of-sample windows. The entire procedure is illustrated in @fig-tourism-cv.

@@

Three patterns stand out. First, the vanilla *MinT-S* (solid green) slightly outperform *MinT-N* (solid purple) at most horizons, although the gap is small. Second, variants that alter the multi-step covariance, either via scaled variance or direct $h$-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this empirical analysis. Most importantly, incorporating a single dominant factor via PC-adjustment (dashed line with round points) delivers best performance among all methods. Both *MinT-S(PC1)* and *MinT-N(PC1)* consistently beat their unadjusted counterparts across horizons. However, adding more than one PC brings no additional benefit and can erode performance, likely due to injecting idiosyncratic noise from weaker components.

@@

@fig-tourism-MSE-heat complements the line plot with a heatmap that standardises each method’s MSE improvement to a 0–100 scale to enhance visual discrimination across horizons. Darker shades indicate larger improvements. The heatmap highlights the pattern from the line plot: *MinT-N(PC1)* and *MinT-S(PC1)* exhibit the highest concentration of dark cells and are repeatedly marked as the best method across horizons (star markers). Overall, the results underscores the consistent gains from PC adjustment and highlights the diminishing returns from more complex covariance treatments (lighter cells for scaled variance and $h$-step residuals).

@@

To dissect the 95% Winkler score results, @fig-tourism-coverage-95-level breaks down performance by hierarchical level, and examines empirical coverage of the 95% prediction intervals. The left panel shows that *MinT-S* produces inferior forecasts compared to the base forecasts at all levels, especially in higher aggregated levels. From our inspection, this is due to the overly shrunk variances from the shrinkage estimator, leading to narrow prediction intervals. The right panel confirms this, as *MinT-S* has the lowest empirical coverage across all levels, indicating that its prediction intervals fail to capture the true observations adequately and resulting in poor Winkler scores. *MinT-N* has relatively good coverage and improved Winkler score at bottom levels, but still is inferior at higher levels. The PC-adjusted variants seem to strike a better balance, improving overall coverage and relative Winkler scores over the base.

@@

1. **ReconCov**: The main *R package* implements the studied covariance estimators, MinT reconciliation, general data generating process functions. Refer to [@ReconCov] for installation and documentation.

>>>>

A collection of time series connected via a set of linear constraints is known as hierarchical time series. Forecasting these series without respecting the hierarchical nature of the data can lead to incoherent forecasts across aggregation levels and, in practice, reduced accuracy. Forecast reconciliation corrects this by adjusting base forecasts to satisfy such constraints. Among modern reconciliation methods, Minimum Trace (MinT) is widely used, however, it requires a good estimate of the forecast error covariance matrix. The current practice uses linear shrinkage towards a diagonal target. Furthermore, the covariance estimate is based on in-sample 1-step-ahead forecast errors, then proportionally scale it to approximate $h$-step-ahead covariance matrix. This raises the question of the method's appropriateness for all real-world applications. We study the shortcomings of current practice and assess the impacts of alternative covariance estimators on MinT, including the NOVELIST estimator (shrinkage towards a soft-thresholded target), PC-adjusted shrinkage (which utilises latent factor structures), and horizon-specific estimators that relax proportional scaling. We evaluate the accuracy of MinT using these covariance estimates for both point and probabilistic reconciled forecasts, and demonstrate their effectiveness and improvements over the shrinkage estimator in a complex, large-hierarchical dataset.

@@

I would like to express my sincere gratitude to my supervisors, Prof. George Athanasopoulos and Dr. Shanika Wickramasuriya, for their invaluable guidance, support, and encouragement throughout my Honours research journey. Their guidance and insights have been instrumental not only in shaping this thesis but also in my personal and academic development, helping me understand what it means to be a researcher. Despite their own commitments, they always made time to assist me whenever I knocked on their office doors. I also thank them for the laughs and light‑hearted moments during our weekly meetings, which made the process enjoyable.

I wish to thank Assoc. Prof. Natalia Bailey for being a wonderful Honours coordinator, always ready to help, providing constructive comments after each milestone, and organising the Honours seminar series throughout the year. She also provided valuable advice when I was initially considering pursuing Honours, and I believe I made the right choice. I am also grateful to Prof. Jiti Gao for his advice regarding my decision to pursue Honours.

I am grateful to my friends and peers, as well as to the professors and lecturers, for their insightful classes throughout my Honours studies. I would also like to thank the Monash University Department of Econometrics and Business Statistics for providing a supportive and stimulating academic environment.

Finally, I would like to thank my family for their unwavering support and encouragement throughout my academic journey, allowing me to focus on my studies and research.

@@

In time series forecasting, aggregation occurs in a variety of settings. For example, Starbucks Corporation operates in many countries, and each country has multiple cities where they have outlets. The sales data is structured *hierarchically*: the top level is the company's total sales, which disaggregates into sales by country, then sales by city within each country, and finally down to sales by individual outlet within each city. As a result, there are over 50,000 sales series across all *aggregation levels*, and decision makers need forecasts at each level to manage inventory and plan marketing strategies effectively. The hierarchy can become even more complex when considering the sales of different product categories (e.g., beverages, food) and the sales of products within each category (e.g., lattes, cappuccinos) at each aggregation level. In this case, the structure is called a *grouped structure*, where the aggregation paths are not unique. Such structures also arise in many other decision-making contexts, from supply chains [@Seaman2022-bb; @Angam2025-od] and energy planning [@Di-Modica2021-ad], to macroeconomics [@El-Gemayel2022-hx; @Li2019-vv] and tourism analysis [@Athanasopoulos2009-lp]. Stakeholders in these settings need forecasts at several aggregation levels to allocate resources and manage risk.

In practice, when forecasts are produced for all series (often called *base forecasts*), they typically violate the aggregation constraints observed in the data (e.g., the sum of all countries' sales forecasts does not equal the total sales forecast). Such forecasts are called *incoherent*. Incoherence undermines downstream decisions that require internal consistency and can degrade forecasting performance. To tackle this problem, forecast reconciliation was introduced. Forecast reconciliation, a post-processing step, utilises the information from the hierarchical structure and data to adjust the initially produced base forecasts, so that the resulting *reconciled forecasts* are *coherent* (i.e., respecting the aggregation constraints). It was first introduced by @Hyndman2011-jv, and later developed by @Van_Erven2015-nx, @Hyndman2016-ic, @Ben-Taieb2019-yx, @Wickramasuriya2019-xq, @Wickramasuriya2020-uk, and others who focus on point forecast reconciliation (more accurately, reconciling the means of predictive distributions). Recognising the importance of probabilistic forecasting, probabilistic forecast reconciliation was studied by @Shang2017-ux, @Jeon2019-wm, and @Ben-Taieb2021-bn, and later formally defined by @Panagiotelis2023-fs. @Athanasopoulos2024-as provide a comprehensive review of the literature on forecast reconciliation.

Among the modern methods, the Min Trace (MinT) approach developed by @Wickramasuriya2019-xq is widely used due to its strong theoretical properties for minimising total reconciled forecast error variance, computational efficiency, and robust empirical performance. MinT is later extended to probabilistic reconciliation by @Wickramasuriya2024-mb, showing that it also minimises the negative log score of the reconciled distribution under Gaussian assumptions. @Wickramasuriya2019-xq also argued that modeling spatial autocorrelations directly from the start would be challenging, as in this case of a large collection of time series. Post-processing reconciliation has the advantage of implicitly modeling this spatial autocorrelation structure, especially true for MinT. MinT is implemented in popular R and Python software ecosystems [@O-Hara-Wild2024-we; @Nixtla].

@@

Meanwhile, there has been substantial progress in high-dimensional covariance estimation that can be leveraged for MinT. Building on shrinkage, @Huang2019-ua proposed NOVELIST, which shrinks toward a soft-thresholded target rather than a diagonal one, improving flexibility to sparse targets. Extending beyond linear shrinkage, @Ledoit2012-ga (and further developed by @Ledoit2020-xt) introduced nonlinear shrinkage that replaces sample eigenvalues with data-driven nonlinear transformations of themselves. In a complementary direction, @Fan2013-jz introduced factor-based estimators that explicitly preserve latent low-rank structure while thresholding the idiosyncratic component. Related thresholding approaches, including hard thresholding [@Bickel2008-nh], generalised thresholding [@Rothman2009-yf], and adaptive thresholding [@Cai2011-jf], have also been developed. This growing toolkit provides multiple pathways to maximise the performance of MinT.

@@

When setting $\boldsymbol{M} = \boldsymbol{I}_n$, the $\boldsymbol{P}$ matrix reduces to the OLS reconciliation $\boldsymbol{S} (\boldsymbol{S}' \boldsymbol{S})^{-1} \boldsymbol{S}'$, which corresponds to an orthogonal projection onto $\mathfrak{s}$. For a general positive definite matrix $\boldsymbol{M}$, $\boldsymbol{P}$ is idempotent but not necessarily symmetric, resulting in an oblique projection. Thus, the choice of $\boldsymbol{M}$ determines the projection direction and, in turn, how disagreements among levels are resolved.

@@

This projection view extends seamlessly from point to probabilistic forecast reconciliation. Given an $h$-step-ahead base forecast density with mean $\hat{\boldsymbol{y}}_{t+h|t}$ and covariance $\boldsymbol{W}_h$, via the linear projection, the resulting reconciled forecast density has mean $\tilde{\boldsymbol{y}}_{t+h|t} = \boldsymbol{P} \hat{\boldsymbol{y}}_{t+h|t}$ and covariance $\boldsymbol{P} \boldsymbol{W}_h \boldsymbol{P}'$. @fig-geometry schematically illustrates this projection. Reconciliation takes the (possibly elliptical) base forecast distribution and "pushes" the entire mass linearly onto the coherent subspace, resulting in the reconciled forecast distribution.

@@

@Wickramasuriya2019-xq showed that by setting $\boldsymbol{M} = \boldsymbol{W}_h = \mathbb{E}( \hat{\boldsymbol{e}}_{t+h|t} \; \hat{\boldsymbol{e}}_{t+h|t}' )$, the covariance matrix of the $h$-step-ahead base forecast errors $\hat{\boldsymbol{e}}_{t+h|t} = \boldsymbol{y}_{t+h} - \hat{\boldsymbol{y}}_{t+h|t}$, we essentially minimise the total variance of the reconciled forecast errors across all series. Equivalently, MinT is the unique linear-unbiased reconciler that minimises the trace of $Var[\boldsymbol{y}_{t+h} - \tilde{\boldsymbol{y}}_{t+h|t}] = \boldsymbol{S} \boldsymbol{G}_h \boldsymbol{W}_h \boldsymbol{G}_h' \boldsymbol{S}'$. This method is thus called Minimum Trace reconciliation. The matrix $\boldsymbol{G}_h$ is thus given by:

@@

where, $\hat{\boldsymbol{R}}_{1,\delta}$ is the thresholded correlation matrix. Each element is regularised by:

@@

On the other hand, the optimal threshold $\hat \delta$ does not have a closed-form solution, and is obtained by rolling-window cross-validation procedure. The idea is to find the threshold $\hat \delta$, with the corresponding $\hat \lambda = \hat \lambda (\hat \delta)$ and $\hat{\boldsymbol{R}}^{N}_{1}(\hat \delta, \hat \lambda)$, that minimises the average out-of-sample 1-step-ahead mean squared reconciled forecast errors over all windows. The formal algorithm is given in Algorithm \ref{alg:novelist_cv}.

@@

\For{$i = v:T-1$}
    \State $j = i - v +1$
    \State $\hat{\boldsymbol{W}}_1 = \frac{1}{v} \sum_{t=j}^{i} \hat{\boldsymbol{e}}_{t} \hat{\boldsymbol{e}}_{t}'$
    \State $\hat{\boldsymbol{D}}_1 = \text{diag}(\hat{\boldsymbol{W}}_1)$
    \State $\hat{\boldsymbol{R}}_1 = \hat{\boldsymbol{D}}_1^{-1/2} \hat{\boldsymbol{W}}_1 \hat{\boldsymbol{D}}_1^{-1/2}$
    \For{$\delta \in \Delta$}
        \State Compute thresholded correlation $\hat{\boldsymbol{R}}_{1,\delta}$ using Equation 2
        \State Compute shrinkage parameter $\hat{\lambda}_{\delta}$ using Equation 3
        \State Compute $\hat{\boldsymbol{R}}^{N}_{1,\delta}$ using Equation 1
        \State $\hat{\boldsymbol{W}}^{N}_{1,\delta} = \hat{\boldsymbol{D}}_1^{1/2} \hat{\boldsymbol{R}}^{N}_{1,\delta} \hat{\boldsymbol{D}}_1^{1/2}$
        \State $\boldsymbol{G} = (\boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{1,\delta} \boldsymbol{S})^{-1} \boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{1,\delta}$
        \State Reconciled forecasts $\tilde{\boldsymbol{y}}_{i+1} = \boldsymbol{S} \boldsymbol{G} \hat{\boldsymbol{y}}_{i+1}$
        \State $\tilde{\boldsymbol{e}}_{i+1, \delta} = \boldsymbol{y}_{i+1} - \tilde{\boldsymbol{y}}_{i+1}$
    \EndFor
\EndFor

\State $MSE_{\delta} = \frac{1}{T-v} \sum_{i=v}^{T-1} (\tilde{\boldsymbol{e}}_{i+1, \delta})^2$ for each $\delta \in \Delta$
\State $\hat{\delta} = \arg\min_{\delta \in \Delta} MSE_{\delta}$

@@

The general design of data-generating process for bottom-level series is a stationary VAR(1) process, with the following structure:

@@

These steps result in a positive covariance matrix $\boldsymbol{\Sigma}^+$ where all the elements are positive. To allow a mixture of positive and negative covariance, we randomly flip the signs of them, but need to maintain the positive definiteness. This can be done by pre- and post-multiplying $\boldsymbol{\Sigma}$ by a random diagonal matrix $\boldsymbol{V}$ with diagonal entries sampled from $\{-1, 1\}$ with equal probability. The final covariance matrix is given by $\boldsymbol{\Sigma} = \boldsymbol{V} \boldsymbol{\Sigma}^+ \boldsymbol{V}$.

For all hierarchies in our experiments, we simulate two panel lengths, $T=50$ ("short") and $T=300$ ("long"), generating extra four observations ($T+1$ to $T+4$) to serve as an out-of-sample test set. We perform 500 Monte Carlo replications for each configuration. In each replication, we fit univariate ARIMA models to the training observations using the automatic AICc minimization algorithm of @Hyndman2008-rf, implemented in the *fabletools* package [@O-Hara-Wild2024-we], generating 1–4-step base forecasts. We then reconcile the base forecasts using different reconciliation methods and evaluate their point and probabilistic forecast accuracy on the test set.

@@

In our main experiments, we examine how MinT combined with the different estimators performs as the hierarchy expands. We generate synthetic data from the VAR(1) framework described earlier, varying the number of bottom‐level series, $n_b$, across two structures: a "small" structure with six groups of six bottom-level series ($n_b = 6$x$6 = 36$), and a "large" configuration with two groups of fifty bottom-level series ($n_b = 2$x$50 = 100$).

@@

The first key observation is that methods with shrinkage slightly outperform those with NOVELIST across all scenarios, although the differences are small. Second, the PC-adjusted variants (using one and two principal components) do not yield improvements over the vanilla versions. This is expected since the synthetic data-generating process does not simulate from strong latent factors. Lastly, the scaled variance and $h$-step-ahead residual approaches do not enhance performance as the forecast horizon increases, suggesting that the proportionality assumption may not be severely violated in this VAR(1) setup.

@@

## Other Data-Generating Processes {#sec-sim-other}

@@

These findings motivate our turn to empirical data in the next section, where latent structural features, regime shifts, and noisy, intermittent series may possibly reveal performance differences.

@@

Forecasting domestic tourism flows is of great economic and policy significance, as policymakers and stakeholders rely on accurate forecasts to make informed decisions regarding resource allocation, infrastructure development, and marketing strategies. The domestic tourism flows in Australia exhibit a natural hierarchical and grouped structure, driven both by geography and by purpose of travel. At the top of this hierarchy lies the national total, which is split into the seven states and territories. Each state is further subdivided into tourism zones, which in turn break down into 77 regions. A complete illustration of this geographic hierarchy appears in Appendix @sec-tourism-supplement. Intersecting this geographic hierarchy is a second dimension: policymakers are also interested in travel motives. This partitions tourism flows into four categories: holiday, business, visiting friends and relatives, and other. Altogether, this yields a grouped structure of 560 series, from the most disaggregated regional-purpose cells up to the full national aggregate. @tbl-tourism-structure depicts this structure.

@@

We quantify tourism demand via "visitor nights", the total number of nights spent by Australians away from home. The data is collected via the National Visitor Survey, managed by Tourism Research Australia, using computer-assisted telephone interviews from nearly 120,000 Australian residents aged 15 years and over [@tourism-au].

The data are monthly time series spanning from January 1998 to December 2016, resulting in 228 observations per series. This gives a challenging high-dimensional forecasting problem with a number of series (560) doubling the number of observations, which is ideal for evaluating reconciliation approaches that rely on high-dimensional covariance estimation. The extreme dimensionality over sample size mirrors many contemporary business problems, for instance, Starbucks Corporation sales. Tourism demand is also economically vital yet highly volatile, with geographical and purpose‑specific patterns creating a realistic stress‑test for reconciliation algorithms.

@@

To assess forecasting performance between models, we adopt a rolling-window cross-validation scheme. Beginning with the first 120 monthly observations (January 1998-December 2005) as the initial training set, we obtain the best-fitted ARIMA model for each of the 560 series via the automatic algorithm by minimising AICc from @Hyndman2008-rf, implemented in the *fabletools* package [@O-Hara-Wild2024-we]. The 1- to 12-step-ahead base forecasts are then generated by these ARIMA models, and then reconciled using multiple approaches. To estimate the NOVELIST and its variants, we implement an extra cross-validation procedure within this training window, as described in Algorithm \ref{alg:novelist_cv}. We then roll the training window forward by one month and refit all models, re-estimate reconciliation variants, and produce another batch of 1- to 12-step-ahead forecasts, repeating until the training set reaches December 2015. In total, this results in 97 out-of-sample windows. The entire procedure is illustrated in @fig-tourism-cv.

@@

Three patterns stand out. First, the vanilla *MinT-S* (solid green) slightly outperforms *MinT-N* (solid purple) at most horizons, although the gap is small. Second, variants that alter the multi-step covariance, either via scaled variance or direct $h$-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this empirical analysis. Most importantly, incorporating a single dominant factor via PC-adjustment (dashed line with round points) delivers the best performance among all methods. Both *MinT-S(PC1)* and *MinT-N(PC1)* consistently beat their unadjusted counterparts across horizons. However, adding more than one PC brings no additional benefit and can erode performance, likely due to injecting idiosyncratic noise from weaker components.

@@

@fig-tourism-MSE-heat complements the line plot with a heatmap that standardises each method’s MSE improvement to a 0–100 scale to enhance visual discrimination across horizons. Darker shades indicate larger improvements. The heatmap highlights the pattern from the line plot: *MinT-N(PC1)* and *MinT-S(PC1)* exhibit the highest concentration of dark cells and are repeatedly marked as the best method across horizons (star markers). Overall, the results underscore the consistent gains from PC adjustment and highlight the diminishing returns from more complex covariance treatments (lighter cells for scaled variance and $h$-step residuals).

@@

To dissect the 95% Winkler score results, @fig-tourism-coverage-95-level breaks down performance by hierarchical level, and examines empirical coverage of the 95% prediction intervals. The left panel shows that *MinT-S* produces inferior forecasts compared to the base forecasts at all levels, especially in higher aggregated levels. From our inspection, this is due to the overly shrunk variances from the shrinkage estimator, leading to narrow prediction intervals. The right panel confirms this, as *MinT-S* has the lowest empirical coverage across all levels, indicating that its prediction intervals fail to capture the true observations adequately and result in poor Winkler scores. *MinT-N* has relatively good coverage and improved Winkler score at bottom levels, but still is inferior at higher levels. The PC-adjusted variants seem to strike a better balance, improving overall coverage and relative Winkler scores over the base.

@@

1. **ReconCov**: The main *R package* implements the studied covariance estimators, MinT reconciliation, general data-generating process functions. Refer to [@ReconCov] for installation and documentation.
